A saturation model for impaired learning with enhanced plasticity: Supplementary material

### 2013-04-08 MHCpaper/MHCsupp.tex
rearranged MHC folder
adjusted spacing
match main text better

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=violet}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
\newcounter{tablerow}
\newcommand{\tablerowautorefname}{row}
\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
%
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=2cm]{cascade.svg}}\label{fig:cascade_model}
 \end{myenuma}
 \end{center}
  \caption[Transition probabilities for different models]{Transition probabilities for different models.
  Potentiation induces transitions indicated by solid red arrows, depression indicated by dashed blue arrows.
  States of strong/weak synaptic weight indicated by red/blue circles.
  States of strong/weak synaptic weight
  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.
  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}) and synaptic weight takes only two values.
  } \label{fig:models}
\end{figure}

@@

\begin{figure}
 \begin{center}
  \aligntop{\includegraphics[height=1.7cm]{pooled.svg}}
 \end{center}
  \caption[The pooled resource model]{The pooled resource model.
  Several two-state synapses share resources that are required for potentiation and depression.
  These resources are depleted as more synapses are potentiated or depressed.
  This pool of synapses can be modelled as one compound synapse.} \label{fig:pooled_model}
\end{figure}

@@

The features of the actual experiments that we'd like to capture are:
\begin{enumerate}
  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the mutant.
  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
  \item After pre-training, gain-increase learning is significantly faster in the mutant than in the wild-type.
  \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
  \item For the mutant, gain-increase learning is significantly faster with pre-training than without it.
\end{enumerate}
These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
We will not worry too much about \autoref{it:gaindown}, as gain-decrease learning uses a different mechanism to gain-increase.
We are only modelling the effect of gain-decrease training on \emph{these} synapses.

@@

We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measuring using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
This should be multiplied by the difference in synaptic weights.
But, as this quantity is constant within a model and doesn't change when comparing wild-type and mutant or with/without pre-training, it can be ignored.

@@

\newcommand{\resultsfig}[3]{\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{#3_learn.eps}}\label{fig:#3_learn}
  \item\aligntop{\includegraphics[width=7cm]{#3_learnS.eps}}\label{fig:#3_learnS}
  \item\label{fig:#3_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_WT.eps}}\label{fig:#3_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_KO.eps}}\label{fig:#3_eq_KO}
                  \end{myenumi}
  \item\label{fig:#3_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_nopre.eps}}\label{fig:#3_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_pre.eps}}\label{fig:#3_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_nopre.eps}}\label{fig:#3_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_pre.eps}}\label{fig:#3_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption[#1]{#1 #2.
  Other parameters can be found in \autoref{tr:#3} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#3_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:#3_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#3_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#3_eq_WT}) wild-type and (\ref{fig:#3_eq_KO}) \KO\ mutant,
  with states of weak synaptic strength to the left and strong states to the right.
  (\ref{fig:#3_pr}) Evolution of probability distributions for (\ref{fig:#3_pr_WT_nopre}\ref{fig:#3_pr_WT_pre}) wild-type and  (\ref{fig:#3_pr_KO_nopre}\ref{fig:#3_pr_KO_pre}) \KO\ mutant without (\ref{fig:#3_pr_WT_nopre},\ref{fig:#3_pr_KO_nopre}) and with (\ref{fig:#3_pr_WT_pre},\ref{fig:#3_pr_KO_pre}) pre-training,
  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#3}
\end{figure}}

\newcommand{\resultsfign}[2]{\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{#2_learn.eps}}\label{fig:#2_learn}
  \item\aligntop{\includegraphics[width=7cm]{#2_learnS.eps}}\label{fig:#2_learnS}
  \item\label{fig:#2_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_WT.eps}}\label{fig:#2_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_KO.eps}}\label{fig:#2_eq_KO}
                  \end{myenumi}
  \item\label{fig:#2_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_nopre.eps}}\label{fig:#2_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_pre.eps}}\label{fig:#2_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_nopre.eps}}\label{fig:#2_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_pre.eps}}\label{fig:#2_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption[#1]{#1.
  Parameters can be found in \autoref{tr:#2} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#2_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:#2_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#2_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#2_eq_WT}) wild-type and (\ref{fig:#2_eq_KO}) \KO\ mutant,
  with states of weak synaptic strength to the left and strong states to the right.
  (\ref{fig:#2_pr}) Evolution of probability distributions for (\ref{fig:#2_pr_WT_nopre}\ref{fig:#2_pr_WT_pre}) wild-type and  (\ref{fig:#2_pr_KO_nopre}\ref{fig:#2_pr_KO_pre}) \KO\ mutant without (\ref{fig:#2_pr_WT_nopre},\ref{fig:#2_pr_KO_nopre}) and with (\ref{fig:#2_pr_WT_pre},\ref{fig:#2_pr_KO_pre}) pre-training,
  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#2}
\end{figure}}

@@

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiWTnopre}
  \Phi = -\frac{2\Delta fq}{M},
\end{equation}
%
where it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:multiWTpre}
\begin{aligned}
  \Phi &= 16(\Delta f)^2q \, \frac{(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1}}
          {(1-2\Delta f)^M - (1+2\Delta f)^M} \\
       &= -\frac{4\Delta fq}{M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS} (black curves).
On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot due to the factor of $(1+2\Delta f)^{M/2-1}$, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.

@@

Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
Without pre-training:
%
\begin{equation}\label{eq:multiKOnopre}
  \Phi = -2\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
\end{equation}
%
This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$.
This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.

@@

With pre-training:
%
\begin{equation}\label{eq:multiKOpre}
\begin{aligned}
  \Phi &= -4\Delta f q \, \frac{(1+2\Delta f) - \beta(1-2\Delta f)}
          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M}   \,
          \brk{\beta (1-2\Delta f) (1+2\Delta f)}^{M/2-1} \\
       &= -4\Delta f q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
Once again, we see that pre-training will speed up learning when $\Delta f$ is small, whereas, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning.

@@

Now consider what would happen if we did not treat gain-increase and gain-decrease symmetrically.
Let us set $f\pot\dec = f\pot\norm - \overline{\Delta f}$.
Then, \eqref{eq:multiKOpre} becomes
%
\begin{equation}\label{eq:multiKOdiffpre}
\begin{aligned}
  \Phi &= -2\prn{\Delta f + \overline{\Delta f}} q \, \frac{(1+2\overline{\Delta f}) - \beta(1-2\overline{\Delta f})}
          {(1+2\overline{\Delta f})^M - \beta^M(1-2\overline{\Delta f})^M}   \,
          \brk{\beta (1-2\overline{\Delta f}) (1+2\overline{\Delta f})}^{M/2-1} \\
       &= -2\prn{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
\end{aligned}
\end{equation}
%
We can then look at the wild-type by setting $\beta=1$:
%
\begin{equation}\label{eq:multiWTdiffpre}
\begin{aligned}
  \Phi &= -8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \,
          \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
          {(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}   \\
       &= -\frac{2\prn{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
\end{aligned}
\end{equation}
%

@@

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:binWTnopre}
  \Phi = -\Delta f q,
\end{equation}
%
where, once again, it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:binWTpre}
\begin{aligned}
  \Phi &= -2\Delta f q.
\end{aligned}
\end{equation}
%
So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
This can be seen in \autoref{fig:binary}\ref{fig:binary_learnS} (black curves).

Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
Without pre-training:
%
\begin{equation}\label{eq:binKOnopre}
  \Phi = -\frac{2\Delta f q}{1+\beta},
\end{equation}
%
which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS} (solid curves).
As discussed below \eqref{eq:multiKOnopre}, the smaller values of has two effects.
Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
The second effect is much smaller for the very short chain considered here.

With pre-training:
%
\begin{equation}\label{eq:binKOpre}
\begin{aligned}
  \Phi &= -4\Delta f q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
          {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
\end{aligned}
\end{equation}
%
As we've already ruled out this model, we won't waste any time analysing this formula.

@@

Secondly, we assumed that there would be equal rates of potentiation and depression in the untrained situation.
However, relaxing this assumption merely changes \eqref{eq:binKOnopre} to
%
\begin{equation}\label{eq:binKOnoprediff}
  \Phi = -\frac{\Delta f q}{f\dep+\beta f\pot}.
\end{equation}
%
This will not change our conclusions either.

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
\newcounter{tablerow}
\newcommand{\tablerowautorefname}{row}
\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
%
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}

@@

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[height=1.7cm]{multistate.svg}}\label{fig:multistate_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=1.7cm]{binary.svg}}\label{fig:binary_model}\hspace{0.5cm}
  \item\aligntop{\includegraphics[height=2.2cm]{cascade.svg}}\label{fig:cascade_model}
 \end{myenuma}
 \end{center}
  \caption[Transition probabilities for different models]{Transition probabilities for different models.
  Potentiation induces transitions indicated by solid red arrows, depression indicated by dashed blue arrows.
  States of strong/weak synaptic weight indicated by red/blue circles.
  States of strong/weak synaptic weight
  (\ref{fig:multistate_model}) In the multistate model the transition probabilities for potentiation/depression are all equal and it is parameterised by these two values.
  (\ref{fig:binary_model}) The two-state model is parameterised by the two transition probabilities.
  (\ref{fig:cascade_model}) In the cascade model, the transition probabilities decay geometrically with a parameter $x$ (see \cite{Fusi2005cascade}) and synaptic weight takes only two values.
  } \label{fig:models}
\end{figure}

@@

\begin{figure}
 \begin{center}
  \aligntop{\includegraphics[height=4cm]{pooled.svg}}
 \end{center}
  \caption[The pooled resource model]{The pooled resource model.
  Several two-state synapses share resources that are required for potentiation and depression.
  These resources are depleted as more synapses are potentiated or depressed.
  This pool of synapses can be modelled as one compound synapse.} \label{fig:pooled_model}
\end{figure}

@@

The features of the actual experiments that we'd like to capture are:
\begin{enumerate}
  \item Without pre-training, gain-increase learning is significantly faster in the wild-type than in the mutant.
  \item For the wild-type, gain-increase learning is significantly faster without pre-training than with it.
  \item For the mutant, gain-increase learning is significantly faster with pre-training than without it.
  \item After pre-training, gain-increase learning is significantly faster in the mutant than in the wild-type.
  \item Gain-decrease pre-training is slightly faster in the wild-type, but not significantly so.\label{it:gaindown}
\end{enumerate}
These questions will not be affected by any output nonlinearity, as long as it is monotonic and fixed.
We will not worry too much about \autoref{it:gaindown}, as gain-decrease learning uses a different mechanism to gain-increase.
We are only modelling the effect of gain-decrease training on \emph{these} synapses.

@@

We will try to gain some analytic insight to some of these models by looking at the slope of the learning curve at the start of gain-increase training.
This is proportional to the net-flux from the states with strong synaptic weight to the weaker states, measured using the transition probabilities for gain-increase but the equilibrium distribution for either untrained or gain-decrease, assuming that pre-training lasts long enough to reach the equilibrium distribution for gain-decrease.
This should be multiplied by the difference in synaptic weights.
But, as this quantity is constant within a model and doesn't change when comparing wild-type and mutant or with/without pre-training, it can be ignored.

@@

\newcommand{\resultsfig}[3]{\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{#3_learn.eps}}\label{fig:#3_learn}
  \item\aligntop{\includegraphics[width=7cm]{#3_learnS.eps}}\label{fig:#3_learnS}
  \\ \vp\vp
  \item\label{fig:#3_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_WT.eps}}\label{fig:#3_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{#3_eq_KO.eps}}\label{fig:#3_eq_KO}
                  \end{myenumi}
  \\ \vp \vp
  \item\label{fig:#3_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_nopre.eps}}\label{fig:#3_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_WT_pre.eps}}\label{fig:#3_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_nopre.eps}}\label{fig:#3_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#3_pr_KO_pre.eps}}\label{fig:#3_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption[#1]{#1 #2.
  Other parameters can be found in \autoref{tr:#3} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#3_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:#3_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#3_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#3_eq_WT}) wild-type and (\ref{fig:#3_eq_KO}) \KO\ mutant,
  with states of weak synaptic strength to the left and strong states to the right.
  (\ref{fig:#3_pr}) Evolution of probability distributions for (\ref{fig:#3_pr_WT_nopre}\ref{fig:#3_pr_WT_pre}) wild-type and  (\ref{fig:#3_pr_KO_nopre}\ref{fig:#3_pr_KO_pre}) \KO\ mutant without (\ref{fig:#3_pr_WT_nopre},\ref{fig:#3_pr_KO_nopre}) and with (\ref{fig:#3_pr_WT_pre},\ref{fig:#3_pr_KO_pre}) pre-training,
  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#3}
\end{figure}}

\newcommand{\resultsfign}[2]{\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\aligntop{\includegraphics[width=7cm]{#2_learn.eps}}\label{fig:#2_learn}
  \item\aligntop{\includegraphics[width=7cm]{#2_learnS.eps}}\label{fig:#2_learnS}
  \\ \vp \vp
  \item\label{fig:#2_eq}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_WT.eps}}\label{fig:#2_eq_WT}
                    \item\aligntop{\includegraphics[width=7cm]{#2_eq_KO.eps}}\label{fig:#2_eq_KO}
                  \end{myenumi}
  \\ \vp \vp
  \item\label{fig:#2_pr}\begin{myenumi}
                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_nopre.eps}}\label{fig:#2_pr_WT_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_WT_pre.eps}}\label{fig:#2_pr_WT_pre}
                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_nopre.eps}}\label{fig:#2_pr_KO_nopre}
                    \item\aligntop{\includegraphics[width=3cm]{#2_pr_KO_pre.eps}}\label{fig:#2_pr_KO_pre}
                  \end{myenumi}
 \end{myenuma}
 \end{center}
  \caption[#1]{#1.
  Parameters can be found in \autoref{tr:#2} of \autoref{tab:params}.
  Time is measured in units of $1/r$, the mean time between plasticity events.
  (\ref{fig:#2_learn}) Learning curves for wild-type and \KO\ mutant with and without pre-training.
  (\ref{fig:#2_learnS}) Learning curves restricted to gain-increase training.
  (\ref{fig:#2_eq}) Equilibrium distributions without training or with gain-increase/decrease training for (\ref{fig:#2_eq_WT}) wild-type and (\ref{fig:#2_eq_KO}) \KO\ mutant,
  with states of weak synaptic strength to the left and strong states to the right.
  (\ref{fig:#2_pr}) Evolution of probability distributions for (\ref{fig:#2_pr_WT_nopre}\ref{fig:#2_pr_WT_pre}) wild-type and  (\ref{fig:#2_pr_KO_nopre}\ref{fig:#2_pr_KO_pre}) \KO\ mutant without (\ref{fig:#2_pr_WT_nopre},\ref{fig:#2_pr_KO_nopre}) and with (\ref{fig:#2_pr_WT_pre},\ref{fig:#2_pr_KO_pre}) pre-training,
  with states of weak synaptic strength at the top and strong states underneath. } \label{fig:#2}
\end{figure}}

@@

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:multiWTnopre}
  \Phi = \frac{2\abs{\Delta f}q}{M},
\end{equation}
%
where it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:multiWTpre}
\begin{aligned}
  \Phi &= 16(\Delta f)^2q \, \frac{(1-2\Delta f)^{M/2-1} (1+2\Delta f)^{M/2-1}}
          {(1-2\Delta f)^M - (1+2\Delta f)^M} \\
       &= \frac{4\abs{\Delta f}q}{M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
So, we see that pre-training will speed up learning when $\Delta f$ is small (in absolute value), as seen in \autoref{fig:multistate_weak}\ref{fig:multistate_weak_learnS} (black curves).
On the other hand, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning a lot due to the factor of $(1+2\Delta f)^{M/2-1}$, as seen in \autoref{fig:multistate_strong}\ref{fig:multistate_strong_learnS}.

@@

Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
Without pre-training:
%
\begin{equation}\label{eq:multiKOnopre}
  \Phi = 2\abs{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M}.
\end{equation}
%
This will be smaller than \eqref{eq:multiWTnopre} if $\beta<\beta^*(M)$.
This function is plotted in \autoref{fig:multistate_star}\ref{fig:multistate_betastar}, where we can see that it approaches 1 rapidly as we increase $M$.

@@

With pre-training:
%
\begin{equation}\label{eq:multiKOpre}
\begin{aligned}
  \Phi &= 4\abs{\Delta f} q \, \frac{(1+2\Delta f) - \beta(1-2\Delta f)}
          {(1+2\Delta f)^M - \beta^M(1-2\Delta f)^M}   \,
          \brk{\beta (1-2\Delta f) (1+2\Delta f)}^{M/2-1} \\
       &= 4\abs{\Delta f} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f)^2.
\end{aligned}
\end{equation}
%
Once again, we see that pre-training will speed up learning when $\Delta f$ is small, whereas, if $\Delta f$ is close to $-\half$, pre-training will initially slow down learning.

@@

Now consider what would happen if we did not treat gain-increase and gain-decrease symmetrically.
Let us set $f\pot\dec = f\pot\norm - \overline{\Delta f}$.
Then, \eqref{eq:multiKOpre} becomes
%
\begin{equation}\label{eq:multiKOdiffpre}
\begin{aligned}
  \Phi &= 2\abs{\Delta f + \overline{\Delta f}} q \, \frac{(1+2\overline{\Delta f}) - \beta(1-2\overline{\Delta f})}
          {(1+2\overline{\Delta f})^M - \beta^M(1-2\overline{\Delta f})^M}   \,
          \brk{\beta (1-2\overline{\Delta f}) (1+2\overline{\Delta f})}^{M/2-1} \\
       &= 2\abs{\Delta f + \overline{\Delta f}} q\,\frac{(1-\beta)\beta^{M/2-1}}{1-\beta^M} + \CO(\Delta f,\overline{\Delta f})^2.
\end{aligned}
\end{equation}
%
We can then look at the wild-type by setting $\beta=1$:
%
\begin{equation}\label{eq:multiWTdiffpre}
\begin{aligned}
  \Phi &= 8\,\overline{\Delta f}\prn{\Delta f + \overline{\Delta f}} q \,
          \frac{(1-2\overline{\Delta f})^{M/2-1} (1+2\overline{\Delta f})^{M/2-1} }
          {\abs{(1+2\overline{\Delta f})^M - (1-2\overline{\Delta f})^M}}   \\
       &= \frac{2\abs{\Delta f + \overline{\Delta f}} q}{M} + \CO(\Delta f,\overline{\Delta f})^2.
\end{aligned}
\end{equation}
%

@@

First, consider the wild-type, for which $q\pot=q\dep=q$.
Without pre-training:
%
\begin{equation}\label{eq:binWTnopre}
  \Phi = \abs{\Delta f} q,
\end{equation}
%
where, once again, it's worth remembering that $\Delta f<0$.
With pre-training:
%
\begin{equation}\label{eq:binWTpre}
\begin{aligned}
  \Phi &= 2\abs{\Delta f} q.
\end{aligned}
\end{equation}
%
So, we see that pre-training will always speed up learning, unlike what is seen in the experiment.
This can be seen in \autoref{fig:binary}\ref{fig:binary_learnS} (black curves).

Now, consider the mutant, for which $q\pot=\beta q\dep=q$, $\beta<1$.
Without pre-training:
%
\begin{equation}\label{eq:binKOnopre}
  \Phi = \frac{2\abs{\Delta f} q}{1+\beta},
\end{equation}
%
which is always larger that \eqref{eq:binWTnopre}, unlike what is seen in the experiment.
This is also seen in \autoref{fig:binary}\ref{fig:binary_learn},\ref{fig:binary_learnS} (solid curves).
As discussed above (below \eqref{eq:multiKOnopre}), the larger value of $q\dep=q/\beta$ has two effects.
Smaller $\beta$ will increase the probability of depression, speeding up learning, but it will also decrease the probability of being ready for depression, slowing down learning.
For the step-multistate model, we argued that the first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ to leading order.
Here $M=2$, so the leading part of the two effects will cancel.
The subleading effects are responsible for the faster learning in the mutant.

With pre-training:
%
\begin{equation}\label{eq:binKOpre}
\begin{aligned}
  \Phi &= 4\abs{\Delta f} q \, \frac{(1-2\Delta f) - \beta(1+2\Delta f)}
          {(1-2\Delta f)^2 - \beta^2(1+2\Delta f)^2}.
\end{aligned}
\end{equation}
%
As we've already ruled out this model, we won't waste any time analysing this formula.

@@

Secondly, we assumed that there would be equal rates of potentiation and depression in the untrained situation.
However, relaxing this assumption merely changes \eqref{eq:binKOnopre} to
%
\begin{equation}\label{eq:binKOnoprediff}
  \Phi = \frac{\abs{\Delta f} q}{f\dep+\beta f\pot}.
\end{equation}
%
This will not change our conclusions either.
